---
title: "MLCourseProject"
author: "Varun"
date: "`r Sys.Date()`"
output: html_document
---

### Course Project- Machine learning.

### 

## **Background**

Using devices such as *Jawbone Up*, *Nike FuelBand*, and *Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement -- a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how *much* of a particular activity they do, but they rarely quantify *how well they do it*. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: <http://groupware.les.inf.puc-rio.br/har> (see the section on the Weight Lifting Exercise Dataset).

### Acknowledgement:

The data for this project come from this source: <http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har>.
[Ugulino, W.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/collaborator.jsf?p1=ugulino); [Cardador, D.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/collaborator.jsf?p1=debora); [Vega, K.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/collaborator.jsf?p1=katia); [Velloso, E.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/collaborator.jsf?p1=evelloso); Milidiu, R.; [Fuks, H.](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/collaborator.jsf?p1=hugo) [**Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements**](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/work.jsf?p1=10335 "Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements"). Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6.\
\
Read more: [http:/groupware.les.inf.puc-rio.br/har#ixzz4TkLUPdxa](http://web.archive.org/groupware.les.inf.puc-rio.br/har#ixzz4TkLUPdxa)

```{r}

```

```{r}
if(!file.exists("data")){
    dir.create("data")
}
fileURL<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
    download.file(fileURL,destfile = "./data/pml-training.csv",method = "curl")
fileURL2<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileURL2,destfile = "./data/pml-testing.csv",method = "curl")
datedoenloaded<-date()
traindata<-read.csv("./data/pml-training.csv")
testdata<-read.csv("./data/pml-testing.csv")
dim(traindata)
dim(testdata)
getwd()
```

### Exploring the data.

### Cleaning the data

### Manipulating the data

### Analyzing the data
